Stateless Model Checking with Data-Race Preemption Points

### 2016-03-24 oopsla-16/evaluation.tex
Improve the prose.

We tested P2s with 6 multithreaded programs:
% from the 410 test suite % XXX: I would like to say this but this is a lie; figure out what else i can say instead
% each tailored to exercise a different part of the P2 project
\mxtest, for locking algorithm correctness,
\tej, a test of thread lifecycle,
\bct~and \paraguay~for condition variables,
\paradise~for semaphores,
% XXX f u latex
and \rwldgr\\for r/w locks.
For \mxtest, \paradise, and \paraguay, we used the {\tt without\_function} command to blacklist {\tt thr\_create}, {\tt thr\_exit}, and {\tt thr\_join},
and for \mxtest~we enabled \landslide's mutex-testing option
(see \sect{\ref{sec:landslide}}).
We tested Pintoses with 3 programs from the class's provided test suite: \prisema, a test of the kernel scheduling algorithm,
\alarmsimul, for the timer sleep routine,
and \waitsimple, for process lifecycle system calls\footnote{
	Some of the Pintoses were partially implemented,
	so each test could only be run on a subset of the 78 submissions; see ``Number tested'' in Table~\ref{tab:drbugs}.
}.
The source code of all 9 test cases is available at
{\em [submitted as supplementary material]}.
For all tests, we used {\tt without\_function} to blacklist PPs on {\tt malloc}'s internal lock.
In total, our evaluation comprises 629 unique tests (pairs of a test program and a Pintos or P2),
% at least 157 of which could expose bugs.
at least 165 of which will expose bugs under one or more MC trials. % QS bugs + bugs where SSSMC found but QS didnt
% TODO: update this number with any new bugs found by PREEMPT_EVERYWHERE

@@

We represented the MC state-of-the-art with 3
%different
configurations of stand-alone \landslide~on the same test suite:
\begin{itemize}
	\item {\bf SSS-MC-DPOR}: Single state space using the maximal PP set from \sect{\ref{sec:initial-pp}}, explored with DPOR \cite{dpor},
	\item {\bf SSS-MC-ICB}: With PPs as above, but instead using ICB \cite{chess-icb} with BPOR \cite{bpor} to find bugs faster, and
	\item {\bf SSS-MC-Shared-Mem}: % TODO: better name?
		Using ICB+BPOR, configured to insert PPs on any shared memory access
		(determined at runtime, excluding threads' accesses to their own stacks),
		which in principle includes all possible data-race PPs.
\end{itemize}
Because parallelizing DPOR/ICB during SSS-MC is an open research problem \cite{parallel-dpor},
we optimistically gave control experiments a linear speedup of 10 hours per test with 1 CPU.
%To match this,
\quicksand~reports the CPU-time spent in addition to the wall-clock time for a direct comparison
(although, with the growing importance of multicore for performance, \quicksand's inherent parallelism is a convenient benefit).
%Finally,
All tests ran on 12-core 3.2 GHz Xeon W3670 machines. %with 12GB of RAM.
}

@@

{\bf Finding new data-race bugs.}
%We quickly pull ahead of SSS-MC, and ultimately conclude with 179\% as many bugs in total.
%The break-even point is at a negligible 90 seconds.
\revision{Compared to the best SSS-MC series, \quicksand~finds}~more bugs within any CPU budget greater than \revision{200}~seconds,
and ultimately concludes 10 CPU-hours with \revision{151}\% as many bugs in total.
\revision{Before the break-even point at 200 seconds,
\quicksand~lags behind the SSS-MC approaches due to additional start-up overhead from its tenfold parallelism.
However, converting SSS-MC's early CPU-time advantage into faster wall-clock performance remains an open research problem \cite{parallel-dpor}.
In Figure~\ref{fig:dowefindbugsfaster}(b), we give \quicksand~full credit for its inherent parallelism,
and with a 10 core allocation, it outperforms SSS-MC for any fixed budget of wall-clock time.}

% TODO CAMREADY
%\revision{
%%{\bf Variance.}
%% TODO CAMREADY: Probably will need at least a 3rd run of dis.
%Because \quicksand~can be nondeterministic in its scheduling of state spaces,
%we ran the QS-DR tests a second time to quantify the variance.
%% TODO CAMREADY
%% We found the maximum area between any two QS-DR curves to be
%We found the area between the two QS-DR curves to be
%% TODO
%{\bf XX\%}
%of the area between QS-DR and SSS-MC-ICB (the next closest),
%% TODO cite
%which we consider an insignificant amount of variance.
%}

@@

{\bf Partial verification.}
When a MC job times out, the user may prefer a brief summary of what parts of the test were verified, rather than writing off all the CPU time as a waste.
%Beyond using the state space estimator to guess the percent coverage,
We know of no approach to quantify the probability
that a bug would be exposed by an untested interleaving,
but \quicksand~at least reports which subsets of PPs resulted in state spaces that did complete in time.
%\quicksand~does not attempt any such quantitative guarantee,
%but can at least report
On 129 tests, the SSS-MC control experiment timed out after 10 hours with no bugs found.
Among these tests, \quicksand~found bugs in 36.
For the other 93, we show the number of state spaces \quicksand~was able to complete in the ``Average tested subset SSes'' column of Table~\ref{tab:drbugs}.
%and the number of unique PPs among them.
%In 6 cases, \quicksand~also failed to complete anything; beyond these,
%between 1 and 253 state spaces were completed for each test
%Between 0 and 233 unique PPs were tested among some completed subsets, with a mean of 27.2 and median of 7.5.
These completions guarantee that, if the test program could expose a bug,
% Justify implicit hypothesis: Sync PPs + DR PPs = all possible relevant PPs.
it would only be found by a new data-race PP not discovered yet, or by a superset combination of PPs not reached.
\revision{
%While obviously not as strong as full verification,
Prior work \cite{sealing} has argued the value of similar {\em compositional testing} when full verification is intractable,
deferring to the user's expertise to judge the value of each subset of PPs verified.
}

@@

>>>>

We tested P2s with 6 multithreaded programs:
% from the 410 test suite % XXX: I would like to say this but this is a lie; figure out what else i can say instead
% each tailored to exercise a different part of the P2 project
\mxtest, for locking algorithm correctness,
\tej, a test of thread lifecycle,
\bct~and \paraguay~for condition variables,
\paradise~for semaphores,
% XXX f u latex
and \rwldgr\\for r/w locks.
For \mxtest, \paradise, and \paraguay, we used the {\tt without\_function} command to blacklist {\tt thr\_create}, {\tt thr\_exit}, and {\tt thr\_join},
and for \mxtest~we enabled \landslide's mutex-testing option
(see \sect{\ref{sec:landslide}}).
We tested Pintoses with 3 programs from the class's provided test suite: \prisema, a test of the kernel scheduling algorithm,
\alarmsimul, for the timer sleep routine,
and \waitsimple, for process lifecycle system calls\footnote{
	Some of the Pintoses were partially implemented,
	so each test could only be run on a subset of the 78 submissions; see ``Number tested'' in Table~\ref{tab:drbugs}.
}.
The source code of all 9 test cases is available at
{\em [submitted as supplementary material]}.
For all tests, we used {\tt without\_function} to blacklist PPs on {\tt malloc}'s internal lock.
In total, our evaluation comprises 629 unique tests (i.e., pairs of a test program and a Pintos or P2),
% at least 157 of which could expose bugs.
at least 173 of which will expose bugs under one or more MC trials. % QS bugs + bugs where SSSMC found but QS didnt

@@

We represented the MC state-of-the-art with 3
%different
configurations of stand-alone \landslide~on the same test suite:
\begin{itemize}
	\item {\bf SSS-MC-DPOR}: Single state space using the maximal PP set from \sect{\ref{sec:initial-pp}}, explored with DPOR \cite{dpor},
	\item {\bf SSS-MC-ICB}: With PPs as above, but instead using ICB \cite{chess-icb} with BPOR \cite{bpor} to find bugs faster, and
	\item {\bf SSS-MC-Shared-Mem}: % TODO: better name?
		Using ICB+BPOR, configured to preempt on any shared memory access \cite{inspect}
		(decided at runtime, excluding threads' accesses to their own stacks),
		which in principle includes all possible data-race PPs.
\end{itemize}
Because parallelizing DPOR/ICB during SSS-MC is an open research problem \cite{parallel-dpor},
we optimistically gave control experiments a linear speedup of 10 hours per test with 1 CPU.
%To match this,
\quicksand~reports the CPU-time spent in addition to the wall-clock time for a resource-fair comparison
(although, with the growing importance of multicore for performance, \quicksand's inherent parallelism is a convenient benefit).
%Finally,
All tests ran on 12-core 3.2 GHz Xeon W3670 machines. %with 12GB of RAM.
}

@@

{\bf Finding new data-race bugs.}
%We quickly pull ahead of SSS-MC, and ultimately conclude with 179\% as many bugs in total.
%The break-even point is at a negligible 90 seconds.
\revision{Compared to the best SSS-MC series, \quicksand~finds}~more bugs within any CPU budget greater than \revision{200}~seconds,
and ultimately concludes 10 CPU-hours with \revision{151}\% as many bugs in total.
\revision{Before the break-even point at 200 seconds,
\quicksand~lags behind the SSS-MC approaches due to additional start-up overhead from its tenfold parallelism.
However, converting SSS-MC's early CPU-time advantage into faster wall-clock performance remains an open research problem \cite{parallel-dpor}.
Figure~\ref{fig:dowefindbugsfaster}(b) gives \quicksand~full credit for its inherent parallelism:
with a 10 core allocation, it outperforms SSS-MC for any fixed budget of wall-clock time.}

% TODO CAMREADY
%\revision{
%%{\bf Variance.}
%% TODO CAMREADY: Probably will need at least a 3rd run of dis.
%Because \quicksand~can be nondeterministic in its scheduling of state spaces,
%we ran the QS-DR tests a second time to quantify the variance.
%% TODO CAMREADY
%% We found the maximum area between any two QS-DR curves to be
%We found the area between the two QS-DR curves to be
%% TODO CAMREADY
%{\bf XX\%}
%of the area between QS-DR and SSS-MC-ICB (the next closest),
%% TODO cite CAMREADY
%which we consider an insignificant amount of variance.
%}

@@

{\bf Partial verification.}
When a MC job times out, the user may prefer a brief summary of what parts of the test were verified, rather than writing off all the CPU time as a waste.
%Beyond using the state space estimator to guess the percent coverage,
%We know of no approach to quantify the probability
%that a bug would be exposed by an untested interleaving,
While recent work \cite{randomized-scheduler} attempts to quantify the probability
that an untested interleaving would expose a bug,
\quicksand~instead reports which subsets of PPs resulted in state spaces that did complete in time.
%\quicksand~does not attempt any such quantitative guarantee,
%but can at least report
% TODO: Fix to be ICB not DPOR.
On 129 tests, SSS-MC-DPOR timed out after 10 hours with no bugs found.
Among these tests, \quicksand~found bugs in 36.
For the other 93, we show the number of state spaces \quicksand~was able to complete in the ``Average tested subset SSes'' column of Table~\ref{tab:drbugs}.
%and the number of unique PPs among them.
%In 6 cases, \quicksand~also failed to complete anything; beyond these,
%between 1 and 253 state spaces were completed for each test
%Between 0 and 233 unique PPs were tested among some completed subsets, with a mean of 27.2 and median of 7.5.
These completions guarantee that, if the test program could expose a bug,
% Justify implicit hypothesis: Sync PPs + DR PPs = all possible relevant PPs.
it would only be found by a new data-race PP not discovered yet, or by a superset combination of PPs not reached.
\revision{
%While obviously not as strong as full verification,
Prior work \cite{sealing} has argued the value of similar {\em compositional testing} when full verification is intractable,
deferring to the user's expertise to judge the value of each subset of PPs verified.
}

@@

% NB. TODO CAMREADY. If a reviewer asks, "How come SSSMC-ShM completed so many state spaces,
% even though last section you argued that so many PPs slows it down terribly?"
% The answer is, "Actually, it's not that SSSMC-ShM magically got fast again for the verification experiment;
% it's that SSSMC-ShM and Quicksand are *BOTH* slow to achieve full completion.
% Quicksand can get bogged down in a combinatorial explosion of different subsets of PPs,
% having a zillion jobs on its hands. Future work should improve this with https://github.com/bblum/landslide/issues/216 ."

### 2016-03-24 oopsla-16/intro.tex
Improve the prose.

On the other hand, when the CPU budget is large enough to fully test all data-race preemption points,
we prove that this constitutes a total verification of all possible thread schedules.
%Short of deciding in advance to preempt on every single instruction \cite{spin}, this was not previously possible without data-race preemption points.
\revision{To achieve the same level of verification, prior model checkers must decide}~in advance to preempt on every single memory access \cite{spin}, \revision{which is computationally prohibitive for
%even moderately-sized
larger tests.
Our approach provides the best of both worlds:
by estimating the size of the test on-the-fly, \quicksand~can find bugs quickly in large tests {\em and} provide fast total verification for small ones.}

We evaluate \quicksand~by testing \numstudence~student thread libraries and kernels from the undergraduate operating systems classes at Carnegie Mellon University, University of California at Berkeley, and University of Chicago.
%We show many advantages over both conventional stateless model checking and single-pass data-race analysis.
We find many of the same bugs found by the conventional approach faster using subsets of preemption points,
and that data-race preemption points quickly expose many new bugs that prior model checkers could not find at all.

>>>>

On the other hand, when the CPU budget is large enough to fully test all data-race preemption points,
we prove that this constitutes a total verification of all possible thread schedules.
%Short of deciding in advance to preempt on every single instruction \cite{spin}, this was not previously possible without data-race preemption points.
\revision{To achieve the same level of verification, prior model checkers must decide}~in advance to preempt on every single memory access \cite{spin}, \revision{which is computationally prohibitive for
even moderately-sized
%larger
tests.
Our approach provides the best of both worlds:
by estimating the size of the test on-the-fly, \quicksand~can find bugs quickly in large tests {\em and} provide fast total verification for small ones.}

We evaluate \quicksand~by testing \numstudence~student thread libraries and kernels from the undergraduate operating systems classes at Carnegie Mellon University, University of California at Berkeley, and University of Chicago.
%We show many advantages over both conventional stateless model checking and single-pass data-race analysis.
%We find many of the same bugs found by the conventional approach faster using subsets of preemption points,
We find that data-race preemption points quickly expose many new bugs that prior model checkers could not find at all,
and that they enable full verification of many more tests than before.
%and that they enable full verification of many racy, but non-failing, tests.

### 2016-03-24 oopsla-16/paper.tex
some tiny things
present sssmc-shmem results, and clean up some todos
Improve the prose.

% TODO: remove all changebars before submission
%\setcounter{changebargrey}{0}
%\newcommand\revision[1]{\begin{samepage}\cbstart#1\cbend\end{samepage}}
%\newcommand\revision[1]{\cbstart#1\cbend}
\newcommand\revision[1]{#1}
\definecolor{purple}{RGB}{192,0,192}
%\newcommand\revision[1]{\hilight{purple}{#1}}

@@

\begin{abstract}
Stateless model checking is a powerful technique for testing concurrent programs,
but suffers from exponential state space explosion when the test input parameters are too large.
Several reduction techniques can mitigate this explosion,
but even after pruning equivalent interleavings, the state space size
%for a fixed set of preemption points
is often intractable.
Most prior tools are limited to preempting only on synchronization APIs,
which reduces the space further, but can miss unsynchronized thread communication bugs.
%
Data race detection, another concurrency testing approach, focuses on suspicious memory access pairs during a single test execution.
It avoids concerns of state space size, but is prone to false positives:
spurious reports of access pairs that cannot be reordered to produce a bug.

@@

We achieve better bug-finding results than either single-pass data-race detection or single-state-space model checking alone,
%We dramatically reduce false-positive data race reports,
%%by verifying the absence of bugs in their associated state spaces
%and find bugs caused by nondeterministic data-races in alternate thread interleavings missed entirely by single-pass analysis.
%Compared to prior work,
%, we find bugs faster in smaller subset state spaces, %% not true, or at least, not by very much
%when the maximal state space is too large to explore
finding new bugs with data-race PPs that could not be exposed by preempting only on synchronization APIs.
By using \numstudence~student operating system implementation projects as our test suite, we also show the potential benefit as a debugging platform in educational settings.
Moreover, when all data-race PPs can be fully tested within the CPU budget, we provide a verification as strong as if every single instruction had been used as a PP.

@@

% TODO: proofs->appeendix
\appendix
\section{Appendix Title}

This is the text of the appendix, if you need one.

@@

Acknowledgments, if needed.

>>>>

%\newcommand\revision[1]{\begin{samepage}\cbstart#1\cbend\end{samepage}}
%\newcommand\revision[1]{\cbstart#1\cbend}
\newcommand\revision[1]{#1}
\definecolor{purple}{RGB}{192,0,192}
%\newcommand\revision[1]{\hilight{purple}{#1}}

@@

\begin{abstract}
Stateless model checking is a powerful technique for testing concurrent programs,
but suffers from exponential state space explosion when the test input parameters are too large.
Several reduction techniques can mitigate this explosion,
but even after pruning equivalent interleavings, the state space size
%for a fixed set of preemption points
is often intractable.
Most prior tools are limited to preempting only on synchronization APIs,
which reduces the space further, but can miss unsynchronized thread communication bugs.
%
Data race detection, another concurrency testing approach, focuses on suspicious memory access pairs during a single test execution.
It avoids concerns of state space size,
%but is prone to false positives:
%spurious reports of access pairs that cannot be reordered to produce a bug.
%but data races do not always correspond to observable failures.
but may report races that do not lead to observable failures,
which jeopardizes a user's willingness to use the analysis.

@@

We achieve better bug-finding results than either single-pass data-race detection or single-state-space model checking alone,
%We dramatically reduce false-positive data race reports,
%%by verifying the absence of bugs in their associated state spaces
%and find bugs caused by nondeterministic data-races in alternate thread interleavings missed entirely by single-pass analysis.
%Compared to prior work,
%, we find bugs faster in smaller subset state spaces, %% not true, or at least, not by very much
%when the maximal state space is too large to explore
finding new bugs with data-race PPs that could not be exposed by preempting only on synchronization APIs.
Moreover, when all data-race PPs can be fully tested within the CPU budget, we provide a verification as strong as if every single instruction had been used as a PP.
Between data-race bugs and verifications, \quicksand~is shown to provide the best of both worlds.
By using \numstudence~student operating system implementation projects as our test suite, we also show the potential benefit as a debugging platform in educational settings.

@@

%\appendix
%\section{Appendix Title}
%
%This is the text of the appendix, if you need one.

@@

%Acknowledgments, if needed.
{\em omitted for double-blind review}

### 2016-03-24 oopsla-16/related.tex
present sssmc-shmem results, and clean up some todos
Improve the prose.

{\bf Preempting at instruction granularity} is a prerequisite for using data-race PPs.
However, the resulting state space explosion demands that any such tool either
choose a small subset of instructions to consider as PPs
or be limited to very small test programs.
%However, every such prior tool we know of has serious drawbacks.
SKI \cite{ski} approaches kernel code by statically choosing a random set of instructions in advance, %offsets from the start of the test,
which is perhaps more similar to
%stress testing or
schedule fuzzing \cite{randomized-scheduler} than to exhaustive state space exploration.
%
SPIN \cite{spin} specializes in verifying synchronization primitive implementations such as RCU \cite{rcu}, which is similar to our \mxtest~experiment,
although it requires code to be written in the PROMELA language.
%However, SPIN is stateful rather than stateless, and explicitly storing visited program states rather than using DPOR limits the size of programs that can be practically tested.
%
Inspect \cite{inspect} instruments source code by inserting wrapper calls around all accesses to potentially-shared data.
It identifies such instructions in advance with an over-approximating static alias analysis,
while \landslide~\cite{landslide} traces the memory locations of accesses at runtime.
Both SPIN and Inspect fix their set of PPs in advance, so could be
%combined with \quicksand~
extended with Iterative Deepening in future work.
%so cannot check implementations directly.

% TODO: Expand on MCR and SATcheck.
{\bf \revision{Reduction}~techniques.} Various improvements or alternatives to DPOR have been developed, such as Dynamic Interface Reduction \cite{demeter},
%Maximal Causality Reduction \cite{mcr},
%DPOR for relaxed memory models \cite{tsopso},
%SAT-directed MC \cite{satcheck},
\revision{and R4 for event-driven applications \cite{r4}}.
These are all compatible with our technique.
%
Recent work \cite{tsopso} has extended DPOR for relaxed memory models \cite{memory-consistency-models},
which we do not yet account for in our proofs (\sect{\ref{sec:soundness}}).
%
SATCheck \cite{satcheck} and Maximal Causality Reduction \cite{mcr} replace DPOR by
%encoding the program's behaviour as a SAT or SMT formula,
%then using a solver to search for schedules guaranteed to expose new behaviour.
using a SAT or SMT solver to search for schedules guaranteed to expose new behaviour.
%These approaches provide improved
They improve
reduction by considering values read and written to identify additional independences,
while DPOR considers only addresses.
They generate new schedules at
%the granularity of each memory access,
memory access granularity,
which Iterative Deepening could allow to be relaxed for large tests in future work.
%
Parrot \cite{parrot} combines MC with a partially-determinizing runtime for further reduction, but still, fewer than half the non-trivial \revision{tests}~in their evaluation could be completed,
\revision{which motivates \quicksand's CPU-budget-oriented approach.}

\revision{{\bf Restricting preemptions.}
Preemption Sealing \cite{sealing} pre\-sents a mechanism similar to the {\tt within\_function} command (\sect{\ref{sec:landslide}}) for users to manually restrict preemptions.
It demonstrates the need to consider subsets of PPs,
as well as developers' willingness to limit a test case's scope so the resulting state space may be fully verified (\sect{\ref{sec:totalverif}}).
%It justifies our assumption that expert
%programmers can customize small test cases which, when fully verified,
%provide a correctness assurance much stronger than the assurance from
%a partially-complete exploration of a huge state space.
Probabilistic Concurrency Testing (PCT) \cite{randomized-scheduler}
is a randomized algorithm that can quantify the probability of uncovering bugs.
PCT is optimized for tests with impossibly large state spaces,
while for more approachable tests,
its incompatibility with DPOR would significantly delay completion.
Future work could use ETAs to heuristically switch between DPOR and PCT}.
Finally, ICB \cite{chess-icb} is most similar to our work,
as both approaches provide a partial verification
%on some subset of interleavings
when full completion is intractable (\sect{\ref{sec:future}}).
However, ICB is limited to a fixed set of PPs,
\revision{cannot estimate remaining time to total verification,
and cannot incorporate data-race PPs not statically coded in advance}.
%and to our knowledge no algorithm has been proposed to dynamically add data-race PPs during a test with ICB.

>>>>

{\bf Preempting at instruction granularity} is a prerequisite for using data-race PPs.
However, the resulting state space explosion demands that any such tool either
choose a small subset of instructions to consider as PPs
or be limited to very small test programs.
%However, every such prior tool we know of has serious drawbacks.
SKI \cite{ski} approaches kernel code by statically choosing a random set of instructions in advance, %offsets from the start of the test,
which is perhaps more similar to
%stress testing or
schedule fuzzing \cite{randomized-scheduler} than to exhaustive state space exploration.
%
SPIN \cite{spin} specializes in verifying synchronization primitive implementations such as RCU \cite{rcu}, which is similar to our \mxtest~experiment,
although it requires code to be written in the PROMELA language.
%However, SPIN is stateful rather than stateless, and explicitly storing visited program states rather than using DPOR limits the size of programs that can be practically tested.
%
Inspect \cite{inspect} instruments source code by
%inserting wrapper calls around
instrumenting
all accesses to potentially-shared data.
It identifies such instructions in advance with an over-approximating static alias analysis,
while \landslide~\cite{landslide} traces the memory locations of accesses at runtime.
Both SPIN and Inspect fix their set of PPs in advance, so could be
%combined with \quicksand~
extended with Iterative Deepening in future work.
%so cannot check implementations directly.

{\bf \revision{Reduction}~techniques.} Various improvements or alternatives to DPOR have been developed, such as Optimal DPOR \cite{optimal-dpor}, Dynamic Interface Reduction \cite{demeter},
%Maximal Causality Reduction \cite{mcr},
%DPOR for relaxed memory models \cite{tsopso},
%SAT-directed MC \cite{satcheck},
\revision{and R4 for event-driven applications \cite{r4}}.
These are all compatible with our technique.
%
Recent work \cite{tsopso} has extended DPOR for relaxed memory models \cite{memory-consistency-models},
which we do not yet account for in our proofs (\sect{\ref{sec:soundness}}).
%
SATCheck \cite{satcheck} and Maximal Causality Reduction \cite{mcr} replace DPOR by
%encoding the program's behaviour as a SAT or SMT formula,
%then using a solver to search for schedules guaranteed to expose new behaviour.
using a SAT or SMT solver to search for schedules guaranteed to expose new behaviour.
%These approaches provide improved
They improve
reduction by considering values read and written to identify additional independences,
while DPOR considers only addresses.
They generate new schedules at
%the granularity of each memory access,
memory access granularity,
which Iterative Deepening could allow to be relaxed for large tests in future work.
%
Parrot \cite{parrot} combines MC with a partially-determinizing runtime for further reduction, but still, fewer than half the non-trivial \revision{tests}~in their evaluation could be completed,
\revision{which motivates \quicksand's CPU-budget-oriented approach.}

\revision{{\bf Restricting preemptions.}
Preemption Sealing \cite{sealing} pre\-sents a mechanism similar to the {\tt within\_function} command (\sect{\ref{sec:landslide}}) for users to manually restrict preemptions.
It demonstrates the need to consider subsets of PPs,
as well as developers' willingness to limit a test case's scope so the resulting state space may be fully verified (\sect{\ref{sec:totalverif}}).
%It justifies our assumption that expert
%programmers can customize small test cases which, when fully verified,
%provide a correctness assurance much stronger than the assurance from
%a partially-complete exploration of a huge state space.
Probabilistic Concurrency Testing (PCT) \cite{randomized-scheduler}
is a randomized algorithm that can quantify the probability of uncovering bugs.
PCT is optimized for tests with impossibly large state spaces,
while for more approachable tests,
its incompatibility with DPOR would significantly delay completion.
Future work could use ETAs to heuristically switch between DPOR and PCT}.
Finally, ICB \cite{chess-icb} is most similar to our work,
as both approaches provide a partial verification
%on some subset of interleavings
when full completion is intractable (\sect{\ref{sec:future}}).
However, ICB is limited to a fixed set of PPs,
\revision{cannot estimate remaining time to total verification,
and can incorporate data-race PPs only when statically coded in advance}.
%and to our knowledge no algorithm has been proposed to dynamically add data-race PPs during a test with ICB.
